---
title: "P8105_hw6_qh2251"
author: "Paulina Han"
date: "2021/11/25"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r}
#raw data
df_raw = read.csv("birthweight.csv", na = c("", "NA", "Unknown"))

#clean version
df_tidy = df_raw %>% 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace),
         ) 

#check missing values  
sum(is.na(df_tidy))

```

The data set contains `r nrow(df_raw)` observations and `r ncol(df_raw)` variables.I converted `babysex`, `frace`, `malform` and `mrace` into factors. There are no missing values in the data set.


```{r}
#fitting model from paper:DOI: 10.1016/j.orcp.2007.09.001
# wtgain:total weight gain
# smoken: average number of cigarettes smoked per day during pregnancy
# parity: number of live births prior to this pregnancy
# babysex: baby’s sex (male = 1, female = 2)
# ppbmi: mother’s pre-pregnancy BMI
# momage: mother’s age at delivery (years)
lm1 = lm(bwt ~ wtgain + smoken + parity + babysex + ppbmi + momage, data = df_tidy)

summary(lm1)

broom::tidy(lm1) %>% 
  knitr::kable(digits = 3)

#plot 
df_tidy %>% 
  add_predictions(lm1) %>% 
  add_residuals(lm1) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = 0.5)+
  labs(
    title = "Predicted vs. Residuals",
    x = "Predicted",
    y = "Residuals"
    ) 
```

To predict the value of birth weight，I include `wtgain`:mother's total weight gain, `smoken`: average number of cigarettes smoked per day during pregnancy , `parity` : number of live births prior to this pregnancy, `babysex` : baby’s sex (male = 1, female = 2), `ppbmi` : mother’s pre-pregnancy BMI, `momage` : mother’s age at delivery (years) in a linear model. I picked out these predictors according to a study done by Kari Johansson in 2007. Here is the [link](https://www.sciencedirect.com/science/article/pii/S1871403X07000622?via%3Dihub) to their interesting paper in predicting child's birth weight.

We can see from the "Predicted vs. Residuals" plot that the average residuals is approximately 0 and the average predicted is around 3000.